Turbulence in compact to giant H~II regions

### 2023-04-24 strucfunc_paper.tex
Javier changes 2023-04-24
Finish revision of first part of M33 regions

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[J. García Vázquez et al.]{
  J. García Vázquez$^{1}$\thanks{
    E-mail: \href{mailto:jgarciav1600@alumno.ipn.mx}{jgarciav1600@alumno.ipn.mx}
  },
  William J. Henney$^{2}$
  and H. O. Castañeda$^{1}$\thanks{Deceased.}
\\
% List of institutions
$^{1}$Escuela Superior de Física y Matemáticas, Instituto Politécnico Nacional, U.P. Adolfo López Mateos, Zacatenco, Ciudad de México, México C.P. 07738\\
$^{2}$Instituto de Radioastronomía y Astrofísica,
Universidad Nacional Autónoma de México,
Antigua Carretera a Pátzcuaro \#8701,\\
Ex-Hda. San José de la Huerta, 
Morelia, Michoacán, México C.P. 58089\\
}

@@

\subsection{NGC~595 and NGC~604 in M33}
\label{sec:m33-ngc}

M33 is the third largest galaxy in the Local Group located at the distance of 820 kpc \citetext{\SI{1}{\arcsecond} = \SI{4.07}{pc} ; \citealp{2015KamKinematics}} and is of morphological type SA(s)cd.
The two brightest \hii{} regions in the galaxy are NGC 604 and NGC 595.

NGC 595 is the second brightest giant \hii{} region in the M33 galaxy.
Using HST/WFPC-2 images \citet{1996AJ....111.1128M} found approximately 250 stars of the OB type, 13 Supergiants and 10 WR, confirming previous candidates.
\citet{1993AJ....105.1400D} use optical photometry to propose 11 WR stars. 
The ionizing output \(Q_0\) is \SI{e50.88}{photons.s^{-1}} (60 times higher compared to the Orion Nebula) and has a luminosity in \ha{} of \SI{e38.95}{erg.s^{-1}}\citep{2002MNRAS.329..481B}.
The mean non-thermal linewidth in \ha{} is \SI{16.8 \pm 0.1}{km.s^{-1}} \citep{lagrois2009multi}.
\citet{lagrois2011} present a kinematic study of the regions using optical emission lines and and radio 21 cm line observations and find that \ha{} shell morphology that is related to the stellar winds of the several massive stars located in its interior.
They proposed Champagne flows at the periphery of the molecular cloud leading to accelerated ionized material.

@@

NGC 604 is the second brightest \hii{} region in the Local Group after 30~Dor.
Using photometry from the HST \citet{1996ApJ...456..174H} identified that the stars are separated into a primary cluster within NGC 604, and a secondary to the north of the region.
In the primary there are \(\sim\)200 OB stars, candidate to red supergiants (RSGs) and the presence of WR stars.
For the northern cluster they find eight luminous stars and no evolved massive stars.
More Recently, \citet{2011MNRAS.411..235E} studied 10 WR stars and propose at least 4 RSGs candidates.
\citet{2012AJ....143...43F} focused on the massive young stellar objects (MYSOs).
The ionizing output \(Q_0\) is \SI{e51.35}{photons.s^{-1}} (175 times higher compared to the Orion Nebula) and has a \ha{} of \SI{e39.42}{erg.s^{-1}} \citep{2002MNRAS.329..481B}.
The mean non-thermal linewidth in \ha{} is \SI{17.8 \pm 0.3}{km.s^{-1}} \citep{1986A&A...160..374H}.
It is the brightest giant \hii{} region in the M33 galaxy and is visually dominated by a predominant big loop, many expanding shells and different size filaments.

>>>>

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[J. García-Vázquez et al.]{
  J. García-Vázquez$^{1}$\thanks{
    E-mail: \href{mailto:jgarciav1600@alumno.ipn.mx}{jgarciav1600@alumno.ipn.mx}
  },
  William J. Henney$^{2}$
  and H. O. Castañeda$^{1}$\thanks{Deceased.}
\\
% List of institutions
$^{1}$Escuela Superior de Física y Matemáticas, Instituto Politécnico Nacional, U.P. Adolfo López Mateos, Zacatenco, Ciudad de México, México C.P. 07738\\
$^{2}$Instituto de Radioastronomía y Astrofísica,
Universidad Nacional Autónoma de México,
Antigua Carretera a Pátzcuaro \#8701,\\
Ex-Hda. San José de la Huerta, 
Morelia, Michoacán, México C.P. 58089\\
}

@@

\startNEW
\subsection{NGC~595 and NGC~604 in M33}
\label{sec:m33-ngc}
\stopNEW

M33 is the third largest galaxy in the Local Group,
located at a distance of \SI{820}{kpc}
\citetext{\SI{1}{\arcsecond} = \SI{4.07}{pc}; \citealp{2015KamKinematics}}
and is of morphological type SA(s)cd.
The two brightest \hii{} regions in the galaxy are NGC~604 and NGC~595.

NGC 595 is the second most luminous giant \hii{} region in the M33 galaxy,
\startNEW
excited by a stellar population comprising
250 OB stars, 13 hot supergiants, and 11 Wolf-Rayet stars \citep{1993AJ....105.1400D, 1996AJ....111.1128M}.
\stopNEW
The ionizing output \(Q_0\) is \SI{e50.88}{photons.s^{-1}} (60 times higher compared to the Orion Nebula) and
the \hii{} region has a luminosity in \ha{} of \SI{e38.95}{erg.s^{-1}} \citep{2002MNRAS.329..481B}.
The mean non-thermal linewidth in \ha{} is
\(\sigma\los = \SI{16.8 \pm 0.1}{km.s^{-1}}\) \citep{lagrois2009multi}.
\citet{lagrois2011} present a kinematic study of the regions using optical emission lines and and radio 21 cm line observations and find an \ha{} shell morphology that is related to the stellar winds of the several massive stars located in its interior.
They proposed champagne flows at the periphery of the molecular cloud leading to accelerated ionized material.

@@

NGC 604 is the
\startNEW
most luminous \hii{} region in M33 and
\stopNEW
the second most luminous in the Local Group after 30~Dor.
\startNEW
The dominant stellar cluster is located in the center of the region and contains 200 OB stars and 4 Wolf-Rayet stars \citep{1996ApJ...456..174H, 2011MNRAS.411..235E}.
The optically visible stars come from a population
with an age of \SI{4}{Myr},
although infrared observations detect the presence of
ongoing massive star formation in the region
\citep{2012AJ....143...43F, 2012ApJ...761....3M}.
\stopNEW
The ionizing output is \(Q_0 = \SI{e51.35}{photons.s^{-1}}\)
(175 times higher than the Orion Nebula) and
the region has an \ha{} luminosity of \SI{e39.42}{erg.s^{-1}} \citep{2002MNRAS.329..481B}.
The mean non-thermal linewidth in \ha{} is
\(\sigma\los = \SI{17.8 \pm 0.3}{km.s^{-1}}\) \citep{1986A&A...160..374H}.
\startNEW
The morphology of the region is strikingly similar to that of 30~Dor
(see Figure~\ref{fig:hii-regions})
\stopNEW
and is visually dominated by a predominant big loop, many expanding shells and different size filaments.
